Last glacial cycle glacier erosion potential in the Alps

### 2020-12-31 articles/erosion/alps-erosion.tex
Draft Alps erosion methods. That was easy.

    Summary of PISM physics and boundary conditions.

@@

    We follow the near-quadratic erosion law of \citep{Herman.etal.2015}.

    We refer to erosion potential as the time-integrated erosion rate.

>>>>

    The ice-sheet model set-up was presented, and the results discussed, in an
    earlier publication \citep{Seguinot.etal.2018}. The simulations use the
    Parallel Ice Sheet Model (PISM, development version~e9d2d1f), an open
    source, finite difference, shallow ice sheet model
    \citep{PISM-authors.2017}. The model includes temperature and water-content
    dependent creep, pseudo-plastic basal sliding accounting for till
    dilatation under high water pressure, bedrock deformation under the ice
    load, and a~positive degree-day (PDD) surface mass balance model. The model
    is initialized with assumed present-day ice thickness and equilibrium
    ice and bedrock temperature at 120\,ka, and ran to the present.
    Climate forcing is provided by a~monthly climatology from interpolated
    observational data \citep[WorldClim;][]{Hijmans.etal.2005} and the European
    Centre for Medium-Range Weather Forecasts Reanalysis Interim
    \citep[ERA-Interim;][]{Dee.etal.2011}, amended with temperature lapse-rate
    corrections from the European Project for Ice Coring in Antarctica
    \citep[EPICA;][] {Jouzel.etal.2007}, and in some cases, time-dependent
    palaeo-precipitation reductions. Lower-resolution, alternative climate
    scenarios are analyzed in the discussion part.

    Physical parametres are listed in the earlier open-access publication
    \citep{Seguinot.etal.2018} and complete PISM set-up is stored in long-term
    archived model output metadata for this \todo{DOI}, and for the earlier
    \todo{DOIs} publications.

@@

    Modelled erosion rates, $\dot{e}$, are calculated from the modelled basal
    velocities, $u_\mathrm{b}$, using a near-quadratic empirical erosion law,
    %
    \begin{equation}
        \dot{e} = K_\mathrm{g} u_\mathrm{b}^l ,
    \end{equation}
    %
    where $K_g = 2.7\times 10^{-7}\,m^{1-l}\,a^{l-1}$ and $l = 2.02$ are
    empirical constants fitted to measured suspended sediment loads and
    reconstructed basal velocities of Franz-Joseph Glacier
    \citep{Herman.etal.2015}. It is important to note that there are both large
    uncertainties around these numbers concerning Franz-Joseph Glacier
    itself \todo{quantify}, and little insurance that this erosion law applies
    elsewhere.

    Unlike landscape evolution model studies \aref, this study does not account
    for the feedback of glacier erosion onto bedrock topography and ice dynamics.
    Neither are other forms of erosion accounted for. Instead, the
    time-integrated erosion rate,
    %
    \begin{equation}
        e =  K_\mathrm{g} \int u_\mathrm{b}^l dt,
    \end{equation}
    %
    is later referred to as the cumulative ``erosion potential''. This erosion
    potential is numerically approximated using a time-step of 10\,a in the
    main simulation, and 100\,a in the alternative climate scenarios runs. It
    can already be noticed from the above formula, that with an exponent, $l$,
    higher than one, the cumulative erosion potential has a greater dependency
    on the basal ice dynamics, $u_\mathrm{b}$, than the timing of glacier
    cover, $dt$.
